---
title: "Threshold Data EDA"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(message = F)
```

```{r}
library(ggplot2)
library(dplyr)
library(FNN)
library(mclust)
library(gridExtra) #for ggplot grid
library(astsa)
library(forecast)
```

BE: before endotoxin, not all horses have this (3, 5, and 6) \newline
presed: first half of endotoxin administration (approx. 30mins but variable) \newline
postsed: second half of endotoxin administration, horses were also sedated (approx. 30 mins) \newline
0_1: 0-1hrs after endotoxin \newline
1_3: 1-3hrs after endotoxin \newline
3_5: 3-5hrs after endotoxin \newline
5_7: 5-7hrs after endotoxin (treatment given in this period) \newline
7_10: 7-10hrs \newline
10_15: 10-15hrs \newline
15_20: 15-20hrs \newline
20_end: 20+hrs if available, not all horses have this

```{r}
#Load Files
suffix = ".csv"
all_time_names = c("Before Endotoxin", "Endotoxin Administration-presed", "Endotoxin-Post Sedation", "0-1 hour", "1-3 hours", "3-5 hours", "5-7 hours", "7-10 hours", "10-15 hours", "15-20 hours", "20 hours to end")
all_time_names_no20 = c("Before Endotoxin", "Endotoxin Administration-presed", "Endotoxin-Post Sedation", "0-1 hour", "1-3 hours", "3-5 hours", "5-7 hours", "7-10 hours", "10-15 hours", "15-20 hours")
all_mains = list("Before Endotoxin", "Pre Sedation", "Post Sedation", "0-1h", "1-3h", "3-5h", "5-7h", "7-10h", "10-15h", "15-20h", "20h+")
all_mains_no20 = list("Before Endotoxin", "Pre Sedation", "Post Sedation", "0-1h", "1-3h", "3-5h", "5-7h", "7-10h", "10-15h", "15-20h")
threshold_suffix = "_1000 threshold"

#H01
h1_prefix = "1000 threshold/horse 1 1000 threshold/horse1_"
h1_ths_rr = list()
for (name in all_time_names) {
  path = paste0(h1_prefix, name, threshold_suffix, suffix)
  data = read.csv(path)
  rr_ts = ts(data$RR_i)
  h1_ths_rr = append(h1_ths_rr, list(rr_ts))
}
h1_all_mains = all_mains

#H02
h2_prefix = "1000 threshold/horse 2 1000 threshold/horse2_"
h2_ths_rr = list()
for (name in all_time_names) {
  path = paste0(h2_prefix, name, threshold_suffix, suffix)
  data = read.csv(path)
  rr_ts = ts(data$RR_i)
  h2_ths_rr = append(h2_ths_rr, list(rr_ts))
}
h2_all_mains = all_mains

#H03 (No Before Endo, No 20+)
h3_prefix = "1000 threshold/horse 3 1000 threshold/horse3_"
h3_middle_names = c("Endotoxin Administration-presed", "Endotoxin-Post Sedation", "0-1 hour", "1-3 hours", "3-5 hours", "5-7 hours", "7-10 hours", "10-15 hours", "15-20 hours")
h3_ths_rr = list()
for (name in h3_middle_names) {
  path = paste0(h3_prefix, name, threshold_suffix, suffix)
  data = read.csv(path)
  rr_ts = ts(data$RR_i)
  h3_ths_rr = append(h3_ths_rr, list(rr_ts))
}
h3_all_mains = list("Pre Sedation", "Post Sedation", "0-1h", "1-3h", "3-5h", "5-7h", "7-10h", "10-15h", "15-20h")

#H04
h4_prefix = "1000 threshold/horse 4 1000 threshold/horse4_"
h4_ths_rr = list()
for (name in all_time_names) {
  path = paste0(h4_prefix, name, threshold_suffix, suffix)
  data = read.csv(path)
  rr_ts = ts(data$RR_i)
  h4_ths_rr = append(h4_ths_rr, list(rr_ts))
}
h4_all_mains = all_mains

#H05 (No Before Endo)
h5_prefix = "1000 threshold/horse 5 1000 threshold/horse5_"
h5_middle_names = c("Endotoxin Administration-presed", "Endotoxin-Post Sedation", "0-1 hour", "1-3 hours", "3-5 hours", "5-7 hours", "7-10 hours", "10-15 hours", "15-20 hours", "20 hours to end")
h5_ths_rr = list()
for (name in h5_middle_names) {
  path = paste0(h5_prefix, name, threshold_suffix, suffix)
  data = read.csv(path)
  rr_ts = ts(data$RR_i)
  h5_ths_rr = append(h5_ths_rr, list(rr_ts))
}
h5_all_mains = list("Pre Sedation", "Post Sedation", "0-1h", "1-3h", "3-5h", "5-7h", "7-10h", "10-15h", "15-20h", "20h+")

#H06 (No Before Endo)
h6_prefix = "1000 threshold/horse 6 1000 threshold/horse6_"
h6_middle_names = c("Endotoxin Administration-presed", "Endotoxin-Post Sedation", "0-1 hour", "1-3 hours", "3-5 hours", "5-7 hours", "7-10 hours", "10-15 hours", "15-20 hours", "20 hours to end")
h6_ths_rr = list()
for (name in h6_middle_names) {
  path = paste0(h6_prefix, name, threshold_suffix, suffix)
  data = read.csv(path)
  rr_ts = ts(data$RR_i)
  h6_ths_rr = append(h6_ths_rr, list(rr_ts))
}
h6_all_mains = list("Pre Sedation", "Post Sedation", "0-1h", "1-3h", "3-5h", "5-7h", "7-10h", "10-15h", "15-20h", "20h+")

#H07
h7_prefix = "1000 threshold/horse 7 1000 threshold/horse7_"
h7_ths_rr = list()
for (name in all_time_names) {
  path = paste0(h7_prefix, name, threshold_suffix, suffix)
  data = read.csv(path)
  rr_ts = ts(data$RR_i)
  h7_ths_rr = append(h7_ths_rr, list(rr_ts))
}
h7_all_mains = all_mains

#H08
h8_prefix = "1000 threshold/horse 8 1000 threshold/horse8_"
h8_ths_rr = list()
for (name in all_time_names) {
  path = paste0(h8_prefix, name, threshold_suffix, suffix)
  data = read.csv(path)
  rr_ts = ts(data$RR_i)
  h8_ths_rr = append(h8_ths_rr, list(rr_ts))
}
h8_all_mains = all_mains

#H09 (No 20+)
h9_prefix = "1000 threshold/horse 9 1000 threshold/horse9_"
h9_ths_rr = list()
for (name in all_time_names_no20) {
  path = paste0(h9_prefix, name, threshold_suffix, suffix)
  data = read.csv(path)
  rr_ts = ts(data$RR_i)
  h9_ths_rr = append(h9_ths_rr, list(rr_ts))
}
h9_all_mains = all_mains_no20

#H10 (No 20+)
h10_prefix = "1000 threshold/horse 10 1000 threshold/horse10_"
h10_ths_rr = list()
for (name in all_time_names_no20) {
  path = paste0(h10_prefix, name, threshold_suffix, suffix)
  data = read.csv(path)
  rr_ts = ts(data$RR_i)
  h10_ths_rr = append(h10_ths_rr, list(rr_ts))
}
h10_all_mains = all_mains_no20

#H11 (No 20+)
h11_prefix = "1000 threshold/horse 11 1000 threshold/horse11_"
h11_ths_rr = list()
for (name in all_time_names_no20) {
  path = paste0(h11_prefix, name, threshold_suffix, suffix)
  data = read.csv(path)
  rr_ts = ts(data$RR_i)
  h11_ths_rr = append(h11_ths_rr, list(rr_ts))
}
h11_all_mains = all_mains_no20

#H12 (No 15-20h, 20+)
h12_prefix = "1000 threshold/horse 12 1000 threshold/horse12_"
h12_middle_names = c("Before Endotoxin", "Endotoxin Administration-presed", "Endotoxin-Post Sedation", "0-1 hour", "1-3 hours", "3-5 hours", "5-7 hours", "7-10 hours", "10-15 hours")
h12_ths_rr = list()
for (name in h12_middle_names) {
  path = paste0(h12_prefix, name, threshold_suffix, suffix)
  data = read.csv(path)
  rr_ts = ts(data$RR_i)
  h12_ths_rr = append(h12_ths_rr, list(rr_ts))
}
h12_all_mains = list("Before Endotoxin", "Pre Sedation", "Post Sedation", "0-1h", "1-3h", "3-5h", "5-7h", "7-10h", "10-15h")

#H13
h13_prefix = "1000 threshold/horse 13 1000 threshold/horse13_"
h13_ths_rr = list()
for (name in all_time_names) {
  path = paste0(h13_prefix, name, threshold_suffix, suffix)
  data = read.csv(path)
  rr_ts = ts(data$RR_i)
  h13_ths_rr = append(h13_ths_rr, list(rr_ts))
}
h13_all_mains = all_mains

ths_all = list(h1_ths_rr, h2_ths_rr, h3_ths_rr, h4_ths_rr, h5_ths_rr, h6_ths_rr, h7_ths_rr, h8_ths_rr, h9_ths_rr, h10_ths_rr, h11_ths_rr, h12_ths_rr, h13_ths_rr)
```

```{r}
# create first row of df
# all_mains_abbr = c("Before", "PreSed", "PostSed", "0-1h", "1-3h", "3-5h", "5-7h", "7-10h", "10-15h", "15-20h", "20h+")
# all_horse_codes = c("H1", "H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10", "H11", "H12", "H13")
# row_names = c()
# for (horse in all_horse_codes) {
#   for (main in all_mains_abbr) {
#     row_names = c(row_names, paste(horse, main, sep = "_"))
#   }
# }
# row_dne = c("H3_Before", "H3_20h+", "H5_Before", "H6_Before", "H9_20h+", "H10_20h+", "H11_20h+", "H12_15-20h", "H12_20h+")
# row_names <- row_names[!row_names %in% row_dne]
```

```{r}
#Summary Stat
total_ts_len = rep(0, 11)
num_ts = rep(0, 11)
rr_summary <- data.frame(matrix(0, nrow = 13, ncol = 11))
max_rr <- data.frame(matrix(0, nrow = 13, ncol = 11))
min_rr <- data.frame(matrix(0, nrow = 13, ncol = 11))

rownames(rr_summary) = c("H1", "H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10", "H11", "H12", "H13")
colnames(rr_summary) = c("Before", "PreSed", "PostSed", "0-1h", "1-3h", "3-5h", "5-7h", "7-10h", "10-15h", "15-20h", "20h+")

counter = 0
for (horse in ths_all) {
  counter = counter + 1
  if (counter %in% c(3, 5, 6)) {
    for (i in 1:length(horse)) {
      total_ts_len[i] = total_ts_len[i] + length(horse[[i]])
      num_ts[i] = num_ts[i] + length(horse[i])
      rr_summary[counter,i+1] = mean(horse[[i]])
      max_rr[counter,i+1] = max(horse[[i]])
      min_rr[counter,i+1] = min(horse[[i]])
    }
  }
  else {
    for (i in 1:length(horse)) {
      total_ts_len[i] = total_ts_len[i] + length(horse[[i]])
      num_ts[i] = num_ts[i] + length(horse[i])
      rr_summary[counter,i] = mean(horse[[i]])
      max_rr[counter,i] = max(horse[[i]])
      min_rr[counter,i] = min(horse[[i]])
    }   
  }
}
total_ts_len / num_ts #average number of time series data points in each duration
rr_summary[rr_summary == 0] <- NA
max_rr[max_rr == 0] <- NA
min_rr[min_rr == 0] <- NA
rr_summary
```

```{r}
#Avg. RR intervals of all horses
matplot(t(rr_summary), type = "l", col = 1:ncol(rr_summary), lty = 1, lwd = 2,
        xlab = "", ylab = "RR Interval",
        main = "All Horses Avg. RR Intervals Over Time", xaxt = "n")
axis(1, at = 1:ncol(rr_summary), labels = colnames(rr_summary), cex.axis = 0.7) 
legend("topright", legend = rownames(rr_summary), col = 1:ncol(rr_summary), lty = 1, lwd = 2, cex = 0.7)
```

```{r, cache = T}
#Store # cluster for all windows
clust_summary <- data.frame(matrix(0, nrow = 13, ncol = 11))
rownames(clust_summary) = c("H1", "H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10", "H11", "H12", "H13")
colnames(clust_summary) = c("Before", "PreSed", "PostSed", "0-1h", "1-3h", "3-5h", "5-7h", "7-10h", "10-15h", "15-20h", "20h+")
counter = 0
for (horse in ths_all) {
  counter = counter + 1
  if (counter %in% c(3, 5, 6)) {
    for (i in 1:length(horse)) {
      df = data.frame(x = horse[[i]][-length(horse[[i]])], y = horse[[i]][-1])
      fit <- Mclust(df, modelNames = "VVE")
      clust_summary[counter,i+1] = fit$G
    }
  }
  else {
    for (i in 1:length(horse)) {
      df = data.frame(x = horse[[i]][-length(horse[[i]])], y = horse[[i]][-1])
      fit <- Mclust(df, modelNames = "VVE")
      clust_summary[counter,i] = fit$G
    }   
  }
}
clust_summary[clust_summary == 0] <- NA
clust_summary
```

```{r}
# store df to csv
rr_summary_vec = na.omit(c(t(rr_summary)))
max_rr_vec = na.omit(c(t(max_rr)))
min_rr_vec = na.omit(c(t(min_rr)))
clust_summary_vec = na.omit(c(t(clust_summary)))
# df = data.frame(window = row_names, mean_RR = rr_summary_vec, num_clust = clust_summary_vec)
# df$mean_RR = rr_summary_vec
# df$num_clust = clust_summary_vec
# csv_file <- "df.csv"
# write.csv(df, file = csv_file)
```

```{r}
#Analysis on control vs. 0-1h
#Percentage decrease from before to 0-1h
rr_decrease = round((rr_summary$Before - rr_summary$`0-1h`)/rr_summary$Before, 2)
rr_decrease
#Average percentage decrease is 34.2%
mean(na.omit(rr_decrease))
```


```{r, cache = T, fig.width=9, fig.height=6, fig.cap="Avg. RR intervals of all horses"}
#Avg. RR intervals of horses 7-13
matplot(t(rr_summary[,]), type = "l", col = "black", lty = 1, lwd = 2,
        xlab = "", ylim = c(780, 2150), ylab = "RR Interval",
        main = "All Horses Mean RR Intervals", xaxt = "n")
axis(1, at = 1:ncol(rr_summary), labels = colnames(rr_summary), cex.axis = 0.7) 

polygon(c(0, 3.5, 3.5, 0), c(2200, 2200, 0, 0), col = rgb(0.5, 0, 0, alpha = 0.3), border = NA)
polygon(c(3.5, 6.5, 6.5, 3.5), c(2200, 2200, 0, 0), col = rgb(0, 0, 0, alpha = 0.3), border = NA)
polygon(c(6.5, 12, 12, 6.5), c(2200, 2200, 0, 0), col = rgb(0, 0.5, 0.5, alpha = 0.3), border = NA)

text(x = 0.9, y = 2100, labels = "Endotoxin Admin", pos = 4, cex = 1.3)
text(x = 3.7, y = 2100, labels = "Severe Symptoms", pos = 4, cex = 1.3)
text(x = 7.3, y = 2100, labels = "Recovery Stage", pos = 4, cex = 1.3)
```

```{r, cache = T, fig.width=9, fig.height=6, fig.cap="Avg. RR intervals of horses 7-13"}
#Avg. RR intervals of horses 7-13
matplot(t(rr_summary[c(7:13),]), type = "l", col = 7:ncol(rr_summary), lty = 1, lwd = 2,
        xlab = "", ylim = c(780, 2150), ylab = "RR Interval",
        main = "Horses 7-13 Mean RR Intervals", xaxt = "n")
axis(1, at = 1:ncol(rr_summary), labels = colnames(rr_summary), cex.axis = 0.7) 
# abline(v = 7, col = "red", lty = 2)
# text(x = 6.8, y = 2000, labels = "Treatment Given", col = "blue", pos = 4)

polygon(c(0, 3.5, 3.5, 0), c(2200, 2200, 0, 0), col = rgb(0.5, 0, 0, alpha = 0.3), border = NA)
polygon(c(3.5, 6.5, 6.5, 3.5), c(2200, 2200, 0, 0), col = rgb(0, 0, 0, alpha = 0.3), border = NA)
polygon(c(6.5, 12, 12, 6.5), c(2200, 2200, 0, 0), col = rgb(0, 0.5, 0.5, alpha = 0.3), border = NA)

text(x = 0.9, y = 2100, labels = "Endotoxin Admin", pos = 4, cex = 1.3)
text(x = 3.7, y = 2100, labels = "Severe Symptoms", pos = 4, cex = 1.3)
text(x = 7.3, y = 2100, labels = "Recovery Stage", pos = 4, cex = 1.3)
legend("topright", legend = rownames(rr_summary)[7:13], col = 7:ncol(rr_summary), lty = 1, lwd = 2, cex = 0.8)
```

```{r, cache = T, fig.cap="Number of Components Assigned"}
#number of components of all horses
matplot(t(clust_summary), type = "l", col = 1:ncol(clust_summary), lty = 1, lwd = 2,
        xlab = "", ylab = "# Components",
        main = "Number of Components Assigned", xaxt = "n")
axis(1, at = 1:ncol(clust_summary), labels = colnames(clust_summary), cex.axis = 0.7) 
legend("topright", legend = rownames(clust_summary), col = 1:ncol(clust_summary), lty = 1, lwd = 2, cex = 0.7)

# individual plots
for (i in 1:length(ths_all)) {
  plot(unlist(clust_summary[i,c(1:6)]), type = "o", col = "blue", xaxt = "n", ylim = c(1,9),
     xlab = "", ylab = "# Components", main = paste("Number of Components of Horse", i))
  axis(1, at = 1:6, labels = colnames(clust_summary)[1:6], cex.axis = 0.7) 
}
```

```{r}
df = read.csv("df.csv", header = TRUE)
# for (i in 1:nrow(df)) {
#   df$Window[i] = unlist(strsplit(df$Horse_Window[i], "_"))[2]
# }
# df = df[,-1]
head(df)

# df$Window_3 = df$Window
# for (i in 1:nrow(df)) {
#   if (df$Window[i] %in% c("Before", "PreSed", "PostSed")) {
#     df$Window_3[i] = "Endotoxin Admin"
#   }
#   else if (df$Window[i] %in% c("0-1h", "1-3h", "3-5h")) {
#     df$Window_3[i] = "Severe Symptoms"
#   }
#   else {
#     df$Window_3[i] = "Recovery Stage"
#   }
# }
# csv_file <- "df.csv"
# write.csv(df, file = csv_file)
```

```{r}
# explore sd1 and sd2
mean = mean(h1_ths_rr[[3]])
sdrr = sqrt( mean(h1_ths_rr[[3]]^2) - mean^2)
delta_rr = h1_ths_rr[[3]][1:length(h1_ths_rr[[3]])-1] - h1_ths_rr[[3]][-1]
sdsd = sqrt( mean(delta_rr^2) - mean(delta_rr)^2)

# gamma_0 = mean(h1_ths_rr[[3]]^2)
# gamma_1 = mean( h1_ths_rr[[3]][1:length(h1_ths_rr[[3]])-1] * h1_ths_rr[[3]][-1])
# phi_0 = mean((h1_ths_rr[[3]] - mean)^2)
# phi_1 = mean( (h1_ths_rr[[3]][1:length(h1_ths_rr[[3]])-1] - mean) * (h1_ths_rr[[3]][-1] - mean))

sd1 = sqrt((sdsd^2)/2)
sd2 = sqrt((sdrr^2)*2 - (sdsd^2)/2)
sd1
sd2
```

```{r}
plot(df$Weight, df$Mean)
abline(a = 0, b = weight_mean_lm$coefficient)
# relationship between body weight and mean RR(both standardized)
weight_mean_lm = lm(Mean ~ Weight-1, data = df)
summary(weight_mean_lm)
weight_mean_lm$coefficient
# mean RR decreases(heart beats faster) as body weight increases (p-value = 0.085)

# relationship between age and mean RR
age_mean_lm = lm(Mean ~ Age, data = df)
summary(age_mean_lm)
# not significant
```

```{r}
# remove the effect of body weight from all mean RR intervals 
# rr_adj = rr/weight
std_wt = df$Weight*weight_mean_lm$coefficient
df$std_weight = std_wt
df$rr_diff = df$Mean - df$std_weight
```


```{r}
# standardize
standardize <- function(x) {
  (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE)
}
for (i in 6:ncol(df)) {
  df[,i] = round(standardize(df[,i]), 3)
}
# standardized_file <- "df_std.csv"
# write.csv(df, file = standardized_file)
```
